---
title: "The Proceedings of the Old Bailey, 1674-1913"
output: html_document
author: "Sebastian Batchford, Noah Vad & Emilie Bærnthsen"
date: "29/05/2025"
---
```{r packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)


```


##  Processing and filtering our data from the Old Bailey
The following script seeks to do the following while creating the data-frames we need for our analysis:

- Filters out data from the years before 1800 which are outside of our scope.

- Renames some of the vectors to something more suitable.

- Collates the numerous defendant ages into age-groups.

```{r csv_script, echo=TRUE}

## 1800-1913 all genders

df_ages_years_all <- read_csv("data/finalproject1800-1913-all.csv") #Reading the data-sheet
df_ages_years_all <- df_ages_years_all %>% rename(year = key) #Renaming "key" value to year.
df_ages_years_all <- filter(df_ages_years_all, year >= 1800) #Removing years before 1800

df_ages_years_all <- df_ages_years_all %>%  #Creating four age-groups
  mutate(age_group_1_21 = `1` + `2` + `3` + `4` + `5` + `6` + `7` + `8` + `9` + `10` + `11` + `12` + `13` + `14` + `15` + `16` + `17` + `18` + `19` + `20` + `21`) %>% 
  mutate(age_group_22_30 = `22` + `23` + `24` + `25` + `26` + `27` + `28` + `29` + `30`) %>% 
  mutate(age_group_31_50 =  `31` + `32` + `33` + `34` + `35` + `36` + `37` + `38` + `39` + `40` + `41` + `42` + `43` + `44` + `45` + `46` + `48` + `49` + `50`) %>% 
  mutate(age_group_51_99 = rowSums(df_ages_years_all[, 53:101]))
df_ages_years_all <- df_ages_years_all[, c("year", "doc_count", "age_group_1_21", "age_group_22_30", "age_group_31_50", "age_group_51_99")]


## 1800-1913 men
df_men <- read_csv("data/finalproject1800-1913-men.csv")
df_men <- df_men %>% rename(year = key)
df_men <- filter(df_men, year >= 1800)

df_men <- df_men %>%
  mutate(age_group_1_21 = `1` + `2` + `3` + `4` + `5` + `6` + `7` + `8` + `9` + `10` + `11` + `12` + `13` + `14` + `15` + `16` + `17` + `18` + `19` + `20` + `21`) %>% 
  mutate(age_group_22_30 = `22` + `23` + `24` + `25` + `26` + `27` + `28` + `29` + `30`) %>% 
  mutate(age_group_31_50 =  `31` + `32` + `33` + `34` + `35` + `36` + `37` + `38` + `39` + `40` + `41` + `42` + `43` + `44` + `45` + `46` + `48` + `49` + `50`) %>% 
  mutate(age_group_51_99 = rowSums(df_men[, 53:101]))
df_men <- df_men[, c("year", "doc_count", "age_group_1_21", "age_group_22_30", "age_group_31_50", "age_group_51_99")]

## 1800-1913 women
df_women <- read_csv("data/finalproject1800-1913-women.csv")
df_women <- df_women %>% rename(year = key)
df_women <- filter(df_women, year >= 1800)

df_women <- df_women %>%
  mutate(age_group_1_21 = `1` + `2` + `3` + `4` + `5` + `6` + `7` + `8` + `9` + `10` + `11` + `12` + `13` + `14` + `15` + `16` + `17` + `18` + `19` + `20` + `21`) %>% 
  mutate(age_group_22_30 = `22` + `23` + `24` + `25` + `26` + `27` + `28` + `29` + `30`) %>% 
  mutate(age_group_31_50 =  `31` + `32` + `33` + `34` + `35` + `36` + `37` + `38` + `39` + `40` + `41` + `42` + `43` + `44` + `45` + `46` + `48` + `49` + `50`) %>% 
  mutate(age_group_51_99 = rowSums(df_women[, 53:101]))
df_women <- df_women[, c("year", "doc_count", "age_group_1_21", "age_group_22_30", "age_group_31_50", "age_group_51_99")]

## Punishment categories
df_punish <- read.csv("data/finalproject1800-1913prisontransportingdeath.csv")
df_punish <- df_punish %>% rename(year = key)
df_punish <- filter(df_punish, year >= 1800)

## Offence categories
df_offence <- read.csv("data/finalproject1800-1913-offencecategory.csv")
df_offence <- df_offence %>% rename(year = key)
df_offence <- filter(df_offence, year >= 1800)

```



## Figures
The following code chunks will creates graphs/figures with our date-frames.


### Figure 1: Death sentences by age group, 1800-1913
This figure shows the number of death sentences in trials from 1800 to 1913 by age group.

```{r figure_age_years_all}


dfplot_ages_years_all <- select(df_ages_years_all, year, age_group_1_21, age_group_22_30, age_group_31_50, age_group_51_99)
dfplot_ages_years_all <- pivot_longer(dfplot_ages_years_all, cols = -year, names_to = "age_group", values_to = "executions")

ggplot(dfplot_ages_years_all, aes(x = year, y = executions, color = age_group)) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.0) +
  scale_y_log10() +
  ggtitle("Figure 1: Death sentences by age group, 1800-1913") +
  xlab("Year") +
  ylab("Number of death sentences (log scale)") +
  scale_x_continuous(breaks = seq(1800, 1913, by = 10), expand = c(0, 0)) +
  theme_minimal() +
  scale_color_manual(values = c("age_group_1_21" = "blue", "age_group_22_30" = "green", "age_group_31_50" = "red", "age_group_51_99" = "yellow"), labels = c("1-21 years", "22-30 years", "31-50 years", "51-99 years")) +
  labs(color = "Age Group")

ggsave(filename = "figures/fig1_executions_by_age_group_all.png", width = 10, height = 6)


```



### Figure 2: Death sentences by age group, 1800-1913 (Males)
This figure shows the number of death sentences in trials with male defendants from 1800 to 1913 by age group.

```{r figure_men}


dfplot_men <- select(df_men, year, age_group_1_21, age_group_22_30, age_group_31_50, age_group_51_99)
dfplot_men <- pivot_longer(dfplot_men, cols = -year, names_to = "age_group", values_to = "executions")

ggplot(dfplot_men, aes(x = year, y = executions, color = age_group)) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.0) +
  scale_y_log10() +
  ggtitle("Figure 2: Death sentences by age group, 1800-1913 (Males)") +
  xlab("Year") +
  ylab("Number of death sentences (log scale)") +
  scale_x_continuous(breaks = seq(1800, 1913, by = 10), expand = c(0, 0)) +
  theme_minimal() +
  scale_color_manual(values = c("age_group_1_21" = "blue", "age_group_22_30" = "green", "age_group_31_50" = "red", "age_group_51_99" = "purple"), labels = c("1-21 years", "22-30 years", "31-50 years", "51-99 years")) +
  labs(color = "Age Group")

ggsave(filename = "figures/fig2_executions_by_age_group_men.png", width = 10, height = 6)



```



### Figure 3: Death sentences by age group, 1800-1913 (Females)
This figure shows the number of death sentences in trials with female defendants from 1800 to 1913 by age group.

```{r figure_women}


dfplot_women <- select(df_women, year, age_group_1_21, age_group_22_30, age_group_31_50, age_group_51_99)
dfplot_women <- pivot_longer(dfplot_women, cols = -year, names_to = "age_group", values_to = "executions")

ggplot(dfplot_women, aes(x = year, y = executions, color = age_group)) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.0) +
  scale_y_log10() +
  ggtitle("Figure 3: Death sentences by age group, 1800-1913 (Females)") +
  xlab("Year") +
  ylab("Number of death sentences (log scale)") +
  scale_x_continuous(breaks = seq(1800, 1913, by = 10), expand = c(0, 0)) +
  theme_minimal() +
  scale_color_manual(values = c("age_group_1_21" = "blue", "age_group_22_30" = "green", "age_group_31_50" = "red", "age_group_51_99" = "purple"), labels = c("1-21 years", "22-30 years", "31-50 years", "51-99 years")) +
  labs(color = "Age Group")

ggsave(filename = "figures/fig3_executions_by_age_group_women.png", width = 10, height = 6)



```



### Figure 4: Punishment sentences in trials, 1800-1913
This figure shows the frequency of punishment categories in trials from 1800 to 1913.

```{r figure_punish}

dfplot_punish <- select(df_punish, year, death, corporal, imprison, miscPunish, noPunish, transport)
dfplot_punish <- pivot_longer(dfplot_punish, cols = -year, names_to = "punish_type", values_to = "punish_amount")


ggplot(dfplot_punish, aes(x = year, y = punish_amount, color = punish_type)) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.0) +
  scale_y_log10() +
  ggtitle("Figure 4: Punishment sentences in trials, 1800-1913") +
  xlab("Year") +
  ylab("Frequency of punishment category (log scale)") +
  scale_x_continuous(breaks = seq(1800, 1913, by = 10), expand = c(0, 0)) +
  theme_minimal() +
  scale_color_manual(
  values = c("death" = "green", "corporal" = "blue", "imprison" = "red", "miscPunish" = "purple", "noPunish" = "orange", "transport" = "yellow"),
  labels = c("death" = "Death", "corporal" = "Corporal", "imprison" = "Imprisonment", "miscPunish" = "Misc. Punishment", "noPunish" = "No Punishment", "transport" = "Transportation")) +
  labs(color = "Punishment Category")

ggsave(filename = "figures/fig4_punishment_types.png", width = 10, height = 6)


```



### Figure 5: Offences in trials with death sentences, 1800-1913
This figure shows the frequency of offence categories in trials with death sentences from 1800 to 1913.

```{r figure_offence}


dfplot_offence <- select(df_offence, year, breakingPeace, deception, kill, miscellaneous, royalOffences, theft, violentTheft, sexual, damage,)
dfplot_offence <- pivot_longer(dfplot_offence, cols = -year, names_to = "offence_category", values_to = "offence_amount")

ggplot(dfplot_offence, aes(x = year, y = offence_amount, color = offence_category)) +
  geom_smooth(se = FALSE, method = "loess", linewidth = 1.0) +
  scale_y_log10() +
  ggtitle("Figure 5: Offences in trials with death sentences, 1800-1913") +
  xlab("Year") +
  ylab("Frequency of offence category (log scale)") +
  scale_x_continuous(breaks = seq(1800, 1913, by = 10), expand = c(0, 0)) +
  theme_minimal() +
  scale_color_manual(
  values = c("breakingPeace" = "green", "deception" = "blue", "kill" = "red", "miscellaneous" = "purple", "royalOffences" = "orange", "theft" = "yellow", "violentTheft" = "hotpink", "sexual" = "cyan", "damage" = "black"),
  labels = c("breakingPeace" = "Breaking Peace", "deception" = "Deception", "kill" = "Killing", "miscellaneous" = "Miscellaneous Offences", "royalOffences" = "Royal Offences", "theft" = "Theft", "violentTheft" = "Violent Theft", "sexual" = "Sexual Offences", "damage" = "Damage to Property")) +
  labs(color = "Offence Category")

ggsave(filename = "figures/fig5_offences_committed.png", width = 10, height = 6)



```



